---
title: "exercise-11_challenge1"
author: "SLCornett"
date: "4/14/2022"
output: 
  html_document: 
    highlight: textmate
    theme: darkly
    toc: TRUE
    toc_float: TRUE
---
# Challenge 1\
```{r preliminaries}
library(tidyverse) #always necessary
library(tidyr)
library(dplyr)
library(ggplot2)
library(ggpubr) 
library(skimr)
library(broom)
library(infer)
```


## One-Factor ANOVA and Inference\
A. Load the AVONET dataset
```{r}
f <- "https://raw.githubusercontent.com/difiore/ada-2022-datasets/main/AVONETdataset1.csv"
d <- read_csv(f, col_names = TRUE)
```


B. Filter the dataset to include only the following variables (19 total): Species1, Family1, Order1, Beak.Length_Culmen, Beak.Width, Beak.Depth, Tarsus.Length, Wing.Length, Tail.Length, Mass, Habitat, Migration, Trophic.Level, Trophic.Niche, Min.Latitude, Max.Latitude, Centroid.Latitude, Primary.Lifestyle, and Range.Size
```{r}
#selecting specific variables
d <- d %>% 
  select(Species1, 
         Family1, 
         Order1, 
         Beak.Length_Culmen, #added 12/4/22
         Beak.Width, 
         Beak.Depth, 
         Tarsus.Length, 
         Wing.Length, 
         Tail.Length, 
         Mass, 
         Habitat, 
         Migration, 
         Trophic.Level, 
         Trophic.Niche, 
         Primary.Lifestyle, #added 12/4/22
         Min.Latitude, 
         Max.Latitude, 
         Centroid.Latitude, 
         Range.Size)
head(d) # heck the selection worked
```


C.  Do a bit of exploratory data analysis with this dataset, e.g., using the {skimr} package.
```{r}
skim(d) # can tell me what's categorical and what's numeric data
```
**Q:** Which of the variables are categorical and which are numeric?\
**A:** CATEGORICAL = Species1, Family1, Order1, Habitat, Trophic.Level, Trophic.Niche, Primary.Lifestyle\
NUMERIC = Beak.Length_Culmen, Beak.Width, Beak.Depth, Tarsus.Length, Wing.Length, Tail.Length, Mass, Migration, Min.Latitude, Max.Latitude, Centroid.Latitude, Range.Size\

### Step 1\
1. Make boxplots of log(**Mass**) in relation to [A] **Trophic.Level** and [B] **Migration** behavior type.\
For each plot, drop _from the visualization_ all species records where the categorical variable of interest is missing from the dataset (**drop.na()**). Also, you will want to convert the variable Migration (which is scored as a number: 1, 2, or 3) from class numeric to either being classified as a factor or as a character (string) variable.
```{r}
class(d$Migration) # checking migration is numerical, but not actually numeric data
# as.factor turns Migration into categorical data, and make a column for logMass
d <- mutate(d, Migration = as.factor(Migration), logMass = log(Mass))
class(d$Migration)# check migration now categorical/factor (could also use as.character())

# can filtered out NA to prevent an extra category for NAs
df_1 <- d %>% # pinned and exclamation mark before is "not"
  drop_na(Trophic.Level) # drop_na Drops rows containing missing values, {tidyr}
# mass v. tropic.level plot
mtl_p <- ggplot(data = df_1, # no NAs 
                   aes(x = Trophic.Level,
                       y = logMass
                       )
                   ) + geom_boxplot() # could also make it a violin plot
mtl_p

# migration as factor
df_2 <- d %>% # pinned and exclamation mark before is "not"
  drop_na(Migration) # drop_na Drops rows containing missing values, {tidyr}
#logMass v. Migration plot
mm_p <- ggplot(data = df_2, #no NAs from this dataframe, not from full dateset
                     aes(x = Migration,
                         y = logMass
                         )
                     ) + geom_boxplot() # makes it a violin plot, could also be a boxplot
mm_p
#only filtered out NAs for the plots, not for the working dataset. 
```


### Step 2\
1. Run linear models using the `lm()` function to look at the relationship between log(**Mass**) and **Trophic.Level** and between log(**Mass**) and **Migration**.
```{r}
# linear model of logMass and Trophic.Level
mtl_m <- lm(data = d, logMass ~ Trophic.Level) # linear model of logMass and Tropic level
summary(mtl_m)

# linear model of logMass and Migration
mm_m <- lm(data = d, logMass ~ Migration) # could also use mutate() to make as.factor(Migration) = Migration 
summary(mm_m)
```


2. Examine the output of the resultant linear models. Is log(**Mass**) associated with either **Trophic.Level** or **Migration** category? That is, in the global test of significance, is the F statistic large enough to reject the null hypothesis of an F value of zero?
```{r}
# ANOVA of the same variables?
# ANOVA of LogMass by Trophic.Level
mtl_a <- aov(data = d, logMass ~ Trophic.Level)
summary(mtl_a)
# ANOVA of LogMass by Migration
mm_a <- aov(data = d, logMass ~ Migration)
summary(mm_a)
```


3. Given the regression coefficients returned for your **Migration** model (mm_m), which Migration categor(ies) are different than the reference level? What level is the reference level?\
**A:** \
Relevel and assess differences among the remaining pair of Migration categories.
```{r}

```


### Step 3 \
1. Conduct a post-hoc **Tukey Significant Differences** test to also evaluate which Migration categories differ “significantly” from one another (<https://difiore.github.io/ada-2022/20-module.html#post-hoc-tests-in-anova>).\
[NOTE: The `TukeyHSD()` function is run on the output of the ANOVA (`aov()`) function.]
```{r}

```


### Step 4 \  
Use a permutation approach to inference to generate a null distribution of F statistic values for the model of log(**Mass**) in relation to **Trophic.Level**. You can do this either by programming your own permutation test (e.g., by shuffling values for the predictor or response variable among observations and calculating an F statistic for each replicate) or by using the {infer} workflow and setting `calculate(stat="F")`.
```{r}
# pulling out the F-statistic (nonparametric)
og_F <- aov(logMass ~ Trophic.Level, data = d) %>% # original F-stat
  tidy() %>%
  filter(term == "Trophic.Level")
og_F # results for the F stat and p-value for the omnibus F test

# Could also do this as a loop
perm_F <- d %>% 
  specify(logMass ~ Trophic.Level) %>%
  hypothesise(null = "independence") %>% # independence must be lowercase
  generate(reps = 1000, # 1000 reps
           type = "permute") %>% # permutation
  calculate(stat = "F") # calc F-stat in permutation
visualise(perm_F) + shade_p_value(obs_stat = original_F$statistic, direction = "greater") # or histogram
# P-value Comparison
p_value <- permuted_F %>% get_p_value(obs_stat = original_F$statistic, direction = "greater")
og_F$p.value
```
