---
title: "exercise-11_challenge2"
author: "SLCornett"
date: "4/14/2022"
output: 
  html_document: 
    highlight: textmate
    theme: darkly
    toc: TRUE
    toc_float: TRUE
---
# Challenge 2\
```{r preliminaries}
library(tidyverse) #always necessary
library(tidyr)
library(dplyr)
library(ggplot2)
library(ggpubr) 
library(skimr)
library(broom)
library(infer)
library(mosaic)
```

## One-Factor ANOVA and Inference\
A. Load the AVONET dataset
```{r}
f <- "https://raw.githubusercontent.com/difiore/ada-2022-datasets/main/AVONETdataset1.csv"
d1 <- read_csv(f, col_names = TRUE)
```


B. Filter the dataset to include only the following variables (19 total): Species1, Family1, Order1, Beak.Length_Culmen, Beak.Width, Beak.Depth, Tarsus.Length, Wing.Length, Tail.Length, Mass, Habitat, Migration, Trophic.Level, Trophic.Niche, Min.Latitude, Max.Latitude, Centroid.Latitude, Primary.Lifestyle, and Range.Size
```{r set up and skim}
#selecting specific variables
d1 <- d1 %>% 
  select(Species1, 
         Family1, 
         Order1, 
         Beak.Length_Culmen, #added 12/4/22
         Beak.Width, 
         Beak.Depth, 
         Tarsus.Length, 
         Wing.Length, 
         Tail.Length, 
         Mass, 
         Habitat, 
         Migration, 
         Trophic.Level, 
         Trophic.Niche, 
         Primary.Lifestyle, #added 12/4/22
         Min.Latitude, 
         Max.Latitude, 
         Centroid.Latitude, 
         Range.Size)
head(d1) # heck the selection worked
skim(d1) # can tell me what's categorical and what's numeric data
```
**Q:** Which of the variables are categorical and which are numeric?\
**A:** CATEGORICAL = Species1, Family1, Order1, Habitat, Trophic.Level, Trophic.Niche, Primary.Lifestyle\
NUMERIC = Beak.Length_Culmen, Beak.Width, Beak.Depth, Tarsus.Length, Wing.Length, Tail.Length, Mass, Migration, Min.Latitude, Max.Latitude, Centroid.Latitude, Range.Size\

### Step 1\
1. Create the following two new variables and add them to AVONET dataset:\
[A] Relative beak length, which you should calculate as the residual of log(**Beak.Length_Culmen**) on log(**Mass**).
```{r Beak Length, logMass mutate new variables}
# adding the log calculations to the dataset
d1 <- d1 %>% mutate(logBL = log(Beak.Length_Culmen), logMass = log(Mass))
# linear model
BL_Mass <- lm(data = d1, logBL ~ logMass)
summary(BL_Mass)
# pulling out the residuals
relative_bl_mass <- BL_Mass$residuals
head(relative_bl_mass) #relative relationship

# put back in the dataset
d1 <- d1 %>% mutate(relative_bl_mass)
```


[B] Relative tarsus length, which you should calculate as the residual of log(**Tarsus.Length**) on log(**Mass**).
```{r Tarsus length mutate new variable}
# adding the log calculations to the dataset
d1 <- d1 %>% mutate(logTL = log(Tarsus.Length))
# linear model
TL_Mass <- lm(data = d1, logTL ~ logMass)
summary(TL_Mass)

#  pulling out the residuals 
relative_tl_mass <- TL_Mass$residuals
head(relative_tl_mass) #relative relationship

# put residuals back in the dataset
d1 <- d1 %>% mutate(relative_tl_mass)
```


### Step 2\
Make a boxplot or violin plot of [A] your new **relative tarsus length** variable in relation to **Primary.Lifestyle** and [B] of your new **relative beak length** variable in relation to **Trophic.Niche**
```{r boxplots}
# A: Tl_Mass residuals vs primary lifestyle
rel.tlm_pl <- ggplot(data = d1,  #tl= tarsus length, m = mass, res= residuals, pl = primary lifestyle
                   aes(x = relative_tl_mass, #the relative tarsus length (vs Mass)
                       y = Primary.Lifestyle # primary lifestyle (pl)
                       )
                   ) + geom_violin() # could also make it a box plot
rel.tlm_pl

# B: BL_Mass residuals vs trophic niche
rel.blm_tn <- ggplot(data = d1, #nl= beak length, m = mass, res= residuals, tn = trophic niche
                   aes(x = relative_bl_mass,#the relative beak length (vs Mass) (blm)
                       y = Trophic.Niche # trophic niche (tn)
                       )
                   ) + geom_boxplot() # could also make it a violin plot
rel.blm_tn
```

### Step 3\
1. Run an ANOVA analyses to look at the association between geographic range size and the variable **Migration**. You should first look at the distribution of the variable **Range.Size** and decide whether and how it might need to be transformed.
```{r Transform Migration and range size, ANOVA range size and migration}
#range size tranformation
hist(d1$Range.Size) #looks funky, thus log transformation
# RS transformation
d1 <- d1 %>% mutate(logRS = log(Range.Size)) # making a new variable of log(range size) in dataset for ease
hist(d1$logRS) #much better.

#migration status transformation
class(d1$Migration) # checking migration is numerical, but not actually numeric data
# as.factor turns Migration into categorical data, and make a column for logMass
d1 <- mutate(d1, Migration = as.factor(Migration))
class(d1$Migration)# check migration now categorical/factor (could also use as.character())

#ANOVA
rs_m_a <- aov(data = d1, logRS ~ Migration) # rs = range size, m = migration
# rs_m_a <- tidy(rs_m_a)
summary(rs_m_a)
```
**Q:** Based on the global model, is **range.size** associated with form of **migration**?\
**A:** yes, based on the global model range size is associate with form of migration significantly (***).\
**Q:** How much of the variance in your measure of range size is associated with Migration style?\
**A:** The F-ratio. 


2. **Q:** Given the regression coefficients returned in output of the [ANOVA (presumably since we did this with a linear model in Challenge 1)] model (rs_m_a), which **Migration** categor(ies) are different than the reference level?\
**A:** All the re-leveled ANOVA models give the same output as the original, unlike the linear models in challege 1, where the reference level affected the model's output, so none of them are different from the original reference level.\
**Q:** What level is the reference level?\
**A:** Presumably all levels (for migration, 1,2, and 3) are the reference level.\
[A] Relevel and assess differences among the remaining pair of **Migration** categories.
```{r releveling ANOVA of migration}
# the original ANOVA model for reference
summary(rs_m_a)
# We can easily relevel() what is the baseline group. The result is very similar, but the sign of  β1 is changed.
# the predictor variable, Migration, is a factor vector instead of a numeric vector, the coefficients are reported and interpreted a bit differently
levels(d1$Migration) 
# Relevel to reference level 1
d1 <- d1 %>% mutate(Migration = relevel(Migration, ref = "1")) 
#re-run ANOVA with defined reference level 1
rs_m_a1 <- aov(data = d1, logRS ~ Migration)
summary(rs_m_a1)

# Relevel to reference level 2
d1 <- d1 %>% mutate(Migration = relevel(Migration, ref = "2")) 
#re-run ANOVA with defined reference level 2
rs_m_a2 <- aov(data = d1, logRS ~ Migration)
summary(rs_m_a2)

# Relevel to reference level 3
d1 <- d1 %>% mutate(Migration = relevel(Migration, ref = "3")) 
#re-run ANOVA with defined reference level 3
rs_m_a3 <- aov(data = d1, logRS ~ Migration)
summary(rs_m_a3)
```


[B] Conduct a post-hoc **Tukey Significant Differences** test to also evaluate which **Migration** categories differ “significantly” from one another.\
(<https://difiore.github.io/ada-2022/20-module.html#post-hoc-tests-in-anova>).\
[NOTE: The `TukeyHSD()` function is run on the output of the ANOVA (`aov()`) function.]
```{r TukeyHSD of ANOVA}
# running the tukey test on ANOVA of Range Size vs Migration
rs_m_a
#post-hoc test
posthoc_rs_m_a <- TukeyHSD(rs_m_a, which = "Migration", ordered = TRUE, conf.level = 0.95)
posthoc_rs_m_a # all levels of migration compared to each other
# plotting the posthoc Tukey test for a nice visual summary
plot(posthoc_rs_m_a, xlim = c(0, 4)) # xlim set to show zero
```


### Step 4\
1. Winnow your data to just consider birds from the Infraorder “Passeriformes” (song birds).
```{r winnow}
passer <- d1 %>% filter(Order1 == "Passeriformes")
```


2. Run separate one-factor ANOVA analyses to look at the association between [1] **relative beak length** and **Primary.Lifestyle** and between [2] **relative beak length** and **Trophic.Level**. In doing so…\
  [A] Make boxplots of response variable by each predictor and by the combination of predictors\
  [B] Run linear models for each predictor separately and interpret the model output\
```{r}
# 1. ANOVA of relative beak length and primary lifestyle of Passeriformes (.a = anova)
pass_rblm_pl.a <- aov(data = passer, relative_bl_mass ~ Primary.Lifestyle) # ANOVA, pass = Passeriformes, r = relative
summary(pass_rblm_pl.a)
# 1A Boxplot of relative beak length by primary lifestyle (.p = plot)
pass_rblm_pl.p <- ggplot(data = passer, # passer only filtered dataframe
                         aes(x = relative_bl_mass,# the relative beak length (vs Mass) (rblm)
                             y = Primary.Lifestyle # pl
                             )
                         ) + geom_violin()
pass_rblm_pl.p
# 1B linear model: relative bleak length vs primary lifestyle of Passeriformes (.lm = linear model)
pass_rblm_pl.lm <- lm(data = passer, relative_bl_mass ~ Primary.Lifestyle)
summary(pass_rblm_pl.lm)
```


[2] Relative beak length vs. 
```{r}
# 2 ANOVA of relative beak length and trophic level of Passeriformes (.a = anova)
pass_rblm_tl.a <- aov(data = passer, relative_bl_mass ~ Trophic.Level) # ANOVA, pass = Passeriformes, r = relative 
summary(pass_rblm_tl.a)

# 2A Boxplot of relative beak length by trophic niche 
pass_rblm_tl.p <- ggplot(data = passer, 
                         aes(x = relative_bl_mass, # the relative beak length (vs Mass) (rblm)
                             y = Trophic.Level # the relative beak length (vs Mass) (rblm)
                             )
                         ) + geom_boxplot()
pass_rblm_tl.p

# 2B linear model: relative beak length vs trophic level  of Passeriformes 
pass_rblm_tl.lm <- lm(data = passer, relative_bl_mass ~ Trophic.Level)
summary(pass_rblm_tl.lm)
```


### Step 5\
Run a two-factor model to look at the association between **relative beak length** and both **Primary.Lifestyle** and **Trophic.Level** among passeriforms.
```{r Passer two way ANOVA}
pass_rblm_pl_tl <- aov(data = passer, relative_bl_mass ~ Primary.Lifestyle + Trophic.Level)
summary(pass_rblm_pl_tl)
```
**Q:** Based on the model output, what would you conclude about how relative beak length is related to these two variables?\
**A:** I conclude there is a significant association between relative beak length and trophic level and primary lifestyle as individual variables.


### Step 6\
Finally, run an additional two-way model with the same dataset and predictors, but adding the possibility of an interaction term. To do this, you should modify your model formula using the colon operator (`:`) to specify the interaction, e.g., **relative beak length ~ Primary.Lifestyle + Trophic.Level + Primary.Lifestyle:Trophic.Level**. 
```{r}
pass_rblm_pl_tl.i <- aov(data = passer, relative_bl_mass ~ Primary.Lifestyle + Trophic.Level + Primary.Lifestyle:Trophic.Level) # .i = interaction
summary(pass_rblm_pl_tl.i)
```
**Q:**Based on the model output, what would you now conclude about how relative beak length is related to these two variables?\
**A:** I conclude there is a significant association between relative beak length and trophic level and primary lifestyle as independent variables AND when they intersect with each other. 


### Step 7\
Use the `interaction.plot()` function to visualize the interaction between **Primary.Lifestyle** and **Trophic.Level** (see Module 20).
```{r}
interaction.plot(
  x.factor = passer$Primary.Lifestyle, # predictor 1
  xlab = "Primary Lifestyle",
  trace.factor = passer$Trophic.Level, # predictor 2
  trace.label = "Trophic Level",
  response = passer$relative_bl_mass, # response variable is relative beak length
  fun = base::mean, # make sure we use {base} version
  ylab = "Relative Beak Length"
)
```


### Additional Steps? \
As noted in Module 20, a general rule of thumb for “equal” variances is to compare the largest and smallest within-grouping level standard deviations and, if this value is less than 2, then it is often reasonable to presume the assumption may not be violated. As a an additional step, check whether **relative beak length** is normally distributed across **Trophical.Level** groups and whether the variance in relative beak size is or is not roughly the same across levels.
```{r}
# rel.blm_tl_hist <- ggplot(data = NULL, aes(x = BL_Mass$model$logBL, y = BL_Mass$residuals)) + geom_point() + facet_wrap(d1$Trophic.Level)
# rel.blm_tl_hist

#rel.blm_tl_hist <- ggplot(data = d1, aes(x = relative_bl_mass)) + geom_histogram() + facet_wrap(d1$Trophic.Level)
#rel.blm_tl_hist
# plot the residuals of beak length and Mass (linear model that gave the relative beak length residuals)
# plot(BL_Mass) # ~ d1$Trophic.Level
```
i tried my best and did not succeed, just as coldplay foretold. 

the end of challenge 2. 